Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities

### 2024-01-10 paper/manuscript.Rmd
Review discussion, figures and legends, describe the bioenergetic model

---
title: Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities
author: "Alain Danet, (Sonia Kéfi, Thomas F Johnson), Andrew P. Beckerman"
date: ""
output:
  officedown::rdocx_document:
    base_format: "bookdown::word_document2"
    number_sections: false
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

First, the stability of communities is often evaluated using temporal
variability the metric of interest (e.g. abundance or biomass), because most
natural communities face persistent perturbations so that other stability
metrics such as resilience and resistance are not always applicable
[@arnoldi_resilience_2016]. Understanding temporal community variability,
measured as the inverse of the coefficient of variation
[@tilman_biodiversity_2006; @roscher_identifying_2011;
@thibaut_understanding_2013; @zhao_biodiversity_2022], requires to scale down
the focus to populations. Community stability can be mathematically partitionned
into average population stability and asynchrony [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022], asynchrony measuring how
much each species fluctuations are compensating each other at the community
level. Population stability and asynchrony depend both on the number of species
and their dominance [@loreau_species_2008; @thibaut_understanding_2013; @zhao_biodiversity_2022],
the interactions among them [@thebault_trophic_2005;
@thebault_relationship_2006; @loreau_species_2008; @danet_species_2021;
@eschenbrenner_diversity_nodate] and on the way that species respond to
environmental perturbations [response diversity, @ives_stability_2000;
@elmqvist_response_2003; @ross_how_2023]. Therefore, accounting for the effects
on community structure on population stability and asynchrony is crucial to
improve our understanding of community stability in face of environmental
change.

Average population stability is decreased by external forcing, such as the
intensity and the frequency of perturbations [@ref]. Species richness can
decrease [@tilman_biodiversity_2006; @loreau_species_2008], increase
[@houlahan_negative_2018], or have little effects on population stability. This
effect of species richness on population stability depends on the evenness of
the community such as dominant species contribute more the average of the
community than surbordinate species. At the core of ecological theory,
population stability depends on internal forcing, here species interactions. The
topogical structure and the distribution of interaction strength have tremendous
effects on population stability, such as strong interaction strength are
generally destabilising [@mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005]. In particular for food-webs, top-down and bottom-up
controls can increase population stability.

@@

---

#Here we deliver inference from modelling of complex biodiverse communities on
#how food-web structure (connectance, average interaction strength, average
#trophic level), response diversity, and environmental stochasticity combine to
#shape temporal stability of community biomass, as well as its partition into
#asynchrony and population stability. We then assess how food-web complexity and
#response diversity shape stability-richness relationships.

@@

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017].
[DESCRIPTION of the  MODEL]

@@

As SAE is further dependant on the evenness of species variance, we
partitioned $SAE$ into $SAE_{even}$ and evenness $EVN$, such as $SAE_{even} =
\sqrt{S}$ and $EVN = SAE / SAE_{even}$.

@@

```{r}
sem_cap <- paste0(
  "Structural Equation Model linking initial model parameters to resulting
  temporal stability of biomass. (a) Arrows display positive (continuous) and
  negative (dashed) standardised effects. Initial parameters were initial
  species richness, initial connectance, Prey-Predator Mass Ratio (PPMR), hill
  exponent, variance of stochastic mortality rates (Env. stochasticity), and
  species response diversity. Only the standardised coefficients whose absolute
  values are superior to 0.1 are displayed. A more complete SEM containing
  further partition of Statistical Averaging Effect and Compensatory effects is
  displayed in Fig. SXX. (b) Partial relationships derived from the SEM, points
  displaying the partial standardized residuals.")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here("report", "figures", "p_sem_tot.pdf"))
```

@@

We found that food-web structure had significant effect on temporal population
stability such as average trophic level and type III functional response (i.e.
hill exponent = 3) increase population stability (resp. $r_\delta$= `r rpt` and
`r rph`, $r_\delta$ being the standardized coefficient, Fig. \@ref(fig:sem)).
In turn, the variance of environmental stochasticity had a strong negative effect
on population stability ($r_\delta$= `r rpe`).

We further found that food-web structure had a strong impact on statistical
averaging effects, which raises from independent species fluctuations weighted
the evenness of their variance. As expected the mathematical relationship (eq.
XX), species richness was a main positive driver of statistical averaging effect
($r_\delta$= `r rsr`, Fig. \@ref(fig:sem)), thereby increasing asynchrony.
Average trophic level and average interaction strength had strong negative
effects on statistical averaging effects (resp. $r_\delta$= `r rst` and
`r rsi`), thereby downplaying the portfolio effect raised from species
richness. In turn, we found that average interaction strength had a small
positive effect on compensatory dynamics ($r_\delta$= `r rci`, Fig.
\@ref(fig:sem)). Species response diversity had a strong positive effect on compensatory dynamics
($r_\delta$= `r rcd`).

@@

Population stability had a positive effect twice higher than asynchrony on the
temporal stability of community (resp. $r_\delta$= `r rscp` and `r rsca`). In
turn, asynchrony was equally driven by statistical averaging effects and
compensatory dynamics (resp. $r_\delta$= `r rac` and `r ras`, Fig.
\@ref(fig:sem)).

```{r}
semtot_cap <- "Total effect of food-web structure on community stability,
asynchrony, and population stability. The total effects were derived from the
structural equation model."
```

```{r semtot, fig.cap = semtot_cap, fig.id = "semtot"}
knitr::include_graphics(here::here("report", "figures", "p_sem_tot_stab.png"))
```

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and response diversity among
species had by far the most important effect on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsdiv`, Fig. \@ref(fig:semtot)). Species richness
had a low overall positive total effect on community stability (`r rtsr`)
because its strong total positive effect on asynchrony (`r rtar`) was downplayed
by its total negative effect on population stability (`r rtpr`) and that the
weight of population stability on community stability is twice higher than the
one of asynchrony. Average trophic level, type III functional response and
predator prey bodysize ratio had positive effects on community stability in
total (resp. `r rtst`, `r rtsh`, `r rtsz`, Fig. \@ref(fig:semtot)), driven by
their positive effects on population stability that were weakly counterbalanced
by their negative effects on asynchrony. In turn, connectance and average
interaction strength had low total negative effects on community stability (resp.
`r rtsc` and `r rtsi`, Fig. \@ref(fig:semtot)). While connectance had negative
effects both on population stability and asynchrony, average interaction
strength had a low total positive effect on population stability counterbalanced
by a stronger negative effect on asynchrony.

@@

)
pred_cap <- paste0("Prediction of the statistical model linking the temporal stability
  of community biomass and species richness. Default parameter values: ", pv_cap, ".") %>%
  str_replace_all("\\s+", " ")
```

@@

Using a more complex linear model (i.e. including interactions), we investigated
more in detail the relationship between community stability and species richness
and how this relation is affected by environmental stochasticity and species
response diversity. We found that environmental stochasticity decrease overall
community stability but has only little effect on the shape of the community
stability-species richness relationship (put some numbers, Fig. \@ref(fig:pred)a
and b). In contrasts, species response diversity both increases overall
community stability and leads to positive community stability-species richness
relationships. Surprisingly, we found that despite average interaction strength
has a small overall effect on community stability, high average interaction
strength change the shape of the community stability-species richness
relationship from positive to negative when response diversity is low (put some
numbers, Fig. \@ref(fig:pred)b).

@@

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, confirming previous results in
empirical fish communities [@danet_species_2021], which contrasts with empirical
data in birds and bats assemblages that reported a equivalent influence of
population stability and asynchrony on community stability
[@olivier_urbanization_2020]. Although the number of empirical and theoretical
studies is still limited to compare the respective contribution of asynchrony and
population stability to community stability in competitive assemblages and
food-webs, we suggest two non-exclusive and complementary explanations that
might drive a lower contribution of asynchrony in food-webs compared to
competitive assemblages. First, the range of bodymasses is much higher in
food-webs, which typically include species that differ by several order of
magnitude [@brose_predator_2019], which in turn creates highly uneven biomass
distribution, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits the raise of high portfolio effects and
compensatory dynamics in species rich communities [@zhao_biodiversity_2022].
Second, the trophic interactions characterising food-webs create strong
interdependence among species, which in turn limits the raise of independent
fluctuations (i.e. portfolio effects) and compensatory dynamics. The question of
which mechanisms and community structure features drive differences in the
contribution of asynchrony and population stability in food-webs and competitive
assemblages is still a open question that might be crucial to accurately
address diagnostics and conservation measures. On top of food-web structure,
we found that environmental stochasticity and environmental correlation among
species was the main drivers of community stability, suggesting that stressors
are the first source of stability. This result is in accordance with previous
studies showing strong effects on environmental stochasticity
[@vasseur_environmental_2007] and the loss of response diversity decrease
community stability in tri-trophic food-webs [@ives_stability_2000;
@thebault_trophic_2005].

We further found that average trophic level increased average population
stability, confirming previous theoreical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_nodate] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators  are themselves more stable due
to lower mortality rates with in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
prey [@shanafelt_stability_2018, we can test that?]. In turn, we found that that
average trophic level decreases strongly asynchrony through a negative effect on
statistical averaging effect. Further analysis showed that average trophic level
limited statistical averaging effect by decreasing the evenness of biomass
variability. It is linked to the model assumption that species of higher trophic
level have a lower overall mortality rates, similarly to previous studies
[@vasseur_environmental_2007], stabilising populations and simultaneously
synchronising them. The fact that Predator Size Mass Ratio displays the same
effect, thereby building food-webs with bigger species having lower mortality
rates, further lends toward this hypothesis. Furthermore, average omnivory was
strongly correlated with average trophic level, suggesting that average omnivory
also stabilise population stability, in line with previous theoretical results
in small modules [@mccann_reevaluating_1997]. Our results further suggest that
higher average trophic level results in lower compensatory dynamics raising from
species interactions. It might be explained by the fact that average trophic
level is strongly associated with average omnivory, which might tamper
synchronized predator-prey oscillations that can raise compensatory dynamics
[@mccann_reevaluating_1997].

We however found no big effect of species richness on population stability,
which has been reported to be negative, positive or neutral in previous studies
[@ruijven_contrasting_2007; @houlahan_negative_2018]. [interpretation?].
However, species richness had a strong positive on asynchrony, by raising
statistical averaging effects. [add blabla about richness asynchrony]. In turn,
we found that species richness magnified the effect of the loss of response
diversity on the compensatory dynamics due to environment. Our interpretation is
that higher species richness synchronizes more species fluctuations due to
environmental mortality when species are not independent because each added
species adds (S - 1, S being richness) positive covariance terms in the
stochastic mortality rates, thereby increasing the overall synchrony in
mortality rates when cross correlation are not null. This assumption that each
species are equally correlated to each other seems unlikely to be true but as
far we know, there is compelling evidence yet for alternative correlation
structures. Connectance, quantifying generalism in food-web, had negative
effects on both asynchrony and population stability. We further found that
higher connectance leaded to less even biomass variability, thereby limiting the
raise of the portfolio effects, and leaded to lower compensatory dynamics due
to species interactions. This results in accordance with empirical finding that
a generalist predator can indeed synchronises the dynamics of a large number of
preys [@raimondo_interspecific_2004]. [No idea about connectance decreasing
population stability]. However, the overall effect of connectance on overall
stability was quite low, similarly to previous empirical findings
[@danet_species_2021], but in apparent disagreement with the predominant role of
connectance on stability in theoretical studies [complexity-stability paradox,
@may_will_1972].

Food-web dynamical features, functional response of type III and average
interaction strength, also had an effect on community stability. Our results
confirms the stabilising effect of the functional response of type III on
populations [@uszko_effects_2017; @brose_allometric_2006]. Compared to type II,
functional response of type III models a decrease of the feeding rates of
consumers when their preys have a low biomass which leads to a less dramatic
fluctuations of prey biomass. Furthermore, we found that functional response of
type III decreased asynchrony by decreasing compensatory dynamics due to species
interaction. This effect might be due to prey switching of consumers when one of
them reach low abundance [@mccann_diversitystability_2000]. Average interaction
strength, as connectance, had overall little effect on community stability.
Average interaction strength was found to decrease asynchrony by decreasing the
evenness of biomass variability, thereby decreasing the portfolio effect,
despite having a small positive effect on compensatory dynamics. In turn, we
surprisingly found that average interaction strength had little effects on
population stability, which contrasts with the importance of this parameter in
complexity-stability relationship [@may_will_1972].

@@

We found that environmental stochasticity does not affect the shape of the
community stability - richness but rather control the overall community
stability (i.e. the intercept of the relation). At the opposite, In turn, a
decrease in response diversity dampens the stability-richness relationship. The
reason is that the strength of environmental stochasticity only affect
population stability and not asynchrony, the latter being the driver of
species-richness relationships [@ives_stability_2000; @thebault_trophic_2005;
@thibaut_understanding_2013]. Average interaction strength and connectance that
were found to have little effects on overall stability, were found to have a
strong effect on stability-richness relationships. This result is in accordance
with the complexity-stability theory that predicts that community stability
(i.e. measured as the dominant eigen value of the jacobian matrix) should
decrease as the community is species diversity, has higher connectance, and
higher interaction strength [@may_will_1972]. However, we found that the
negative effects of complexity on stability-richness relationship are cancelled
when response diversity is high. A interpretation is that environmental
stochasticity itself does not override the negative complexity-stability
relationships, but it the presence of response diversity and resulting
asynchrony in species fluctuations which overrides the negative effects of
complexity on stability-richness relationship. Response diversity might then one
mechanism that explains why so much positive relationship between stability and
species richness are found in empirical settings while food-web theoretical
model that do not often include environmental stochasticity and response
diversity finds negative complexity stability relationships
[@ives_stability_2007].

>>>>

---
title: Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities
author: "Alain Danet, (Sonia Kéfi, Thomas F Johnson), Andrew P. Beckerman"
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    base_format: "bookdown::word_document2"
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

First, the stability of communities is often evaluated using temporal
variability the metric of interest (e.g. abundance or biomass), because most
natural communities face persistent perturbations so that other stability
metrics such as resilience and resistance are not always applicable
[@arnoldi_resilience_2016]. Understanding temporal community variability,
measured as the inverse of the coefficient of variation
[@tilman_biodiversity_2006; @roscher_identifying_2011;
@thibaut_understanding_2013; @zhao_biodiversity_2022], requires to scale down
the focus to populations. Community stability can be mathematically partitionned
into average population stability and asynchrony [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022], asynchrony measuring how
much each species fluctuations are compensating each other at the community
level. Population stability and asynchrony depend both on the number of species
and their dominance [@loreau_species_2008; @thibaut_understanding_2013; @zhao_biodiversity_2022],
the interactions among them [@thebault_trophic_2005;
@thebault_relationship_2006; @loreau_species_2008; @danet_species_2021;
@eschenbrenner_diversity_2023] and on the way that species respond to
environmental perturbations [response diversity, @ives_stability_2000;
@elmqvist_response_2003; @ross_how_2023]. Therefore, accounting for the effects
on community structure on population stability and asynchrony is crucial to
improve our understanding of community stability in face of environmental
change.

Average population stability is decreased by external forcing, such as the
intensity and the frequency of perturbations [@loreau_species_2008]. Species richness can
decrease [@tilman_biodiversity_2006; @loreau_species_2008], increase
[@houlahan_negative_2018], or have little effects on population stability. This
effect of species richness on population stability depends on the evenness of
the community such as dominant species contribute more the average of the
community than surbordinate species. At the core of ecological theory,
population stability depends on internal forcing, here species interactions. The
topogical structure and the distribution of interaction strength have tremendous
effects on population stability, such as strong interaction strength are
generally destabilising [@mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005]. In particular for food-webs, top-down and bottom-up
controls can increase population stability.

@@

---
#Here we deliver inference from modelling of complex biodiverse communities on
#how food-web structure (connectance, average interaction strength, average
#trophic level), response diversity, and environmental stochasticity combine to
#shape temporal stability of community biomass, as well as its partition into
#asynchrony and population stability. We then assess how food-web complexity and
#response diversity shape stability-richness relationships.

@@

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017].
This model is widely used to simulate the dynamics of complex ecological
communities because of the simplicity of its parametrisation and that its
parameters are rooted in the metabolic theory of ecology [@brown_toward_2004].
The model describes species biomass dynamics over time with two master equations (eq. \@ref(eq:dbdt)), one for the
primary producers (eq. \@ref(eq:prod)) and one for the consumers (eq. \@ref(eq:prod)):

\begin{subequations}
    \label{eq:dbdt}

    \begin{equation}
        \label{eq:cons}
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i - d_i B_i
    \end{equation}

    \begin{equation}
    	\label{eq:prod}
    	\frac{dB_i}{dt} =
	r_i B_i G_i
	- \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - d_i B_i
    \end{equation}
\end{subequations}

The consumers gain biomass by consumming resources (i.e. their preys) at a rate
that depends on their mass specific metabolic rate ($x_i$),  maximum cumsumption
rate ($y_i$) and on the functional response ($F_ij$) describing how the rate of
consumption of a consumer $i$ on a resource $j$ vary with the biomass of this resource.
Consumers then loses biomass by being consumed, $e_ij$ being the assimilisation
efficiency of the consumer $j$ on the resource $i$. Consumers also lose biomass
over time through metabolic losses ($-x_iB_i$) and by natural mortality due to
environment ($-d_iB_i$). The primary producers gain biomass over time with a
growth rate ($r_i$, eq. \@ref(eq:prod)) and a functional response ($G_i$)
describing how the growth of the producers vary with their biomass. They lose
biomass by being consumed by consumers and by natural mortality.

\begin{subequations}
    \label{eq:func}
    \begin{equation}
        \label{eq:growth}
        G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
    \end{equation}

    \begin{equation}
    	\label{eq:holling}
        F_{ij} = \frac{\omega_{ij} B_j^h}
        {B_0^h + c_i B_i
        	+ \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
    \end{equation}
\end{subequations}

The growth function of the primary producers (eq. \@ref(eq:growth)) has the shape of a logistic growth
where the growth rate is maximal and null when
$\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j$ are respectively close to 0 and 1. $\alpha_{ij}$ is the
per capita effect of the producer $j$ on the producer $i$, $K_i$ being the
carrying capacity for the producer $j$. The functional response of a consumer
feeding on a resource (eq. \@ref(eq:holling)) depends on the relative preference
of the consumer on the resource ($\omega_ij$) is limited by an half-saturation
rate ($B_0$), intraspecific interference coefficient ($c_i$), and on the
availability of its other resources
($\sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h$). Finally, the $h$ exponent
controls the shape of the functional response from a saturating function
(Holling type II) to a sigmoid (Holling type III).

@@

As SAE is further dependant on the evenness of species variance, we
partitioned $SAE$ into $SAE_{even}$ and evenness $EVN$, such as  $EVN = SAE / SAE_{even}$.
$SAE_{even}$ corresponds to the $SAE$ when species quantities are perfectly
even such as from eq. \@ref(eq:sae2), we can obtain $SAE_{even}$ from which $EVN$ can be computed:

\begin{equation}
\begin{aligned}
SAE &= \dfrac{\sum_i \sigma_i}{\sqrt{\sum_i \sigma^2_i}} = \sqrt{\dfrac{1}{\sum_i (\sigma_i\sigma^2_i})^2} = \sqrt{\dfrac{1}{\sum_i (p_i)^2}}\\
SAE_{even} &= \sqrt{\dfrac{1}{\sum_i (1/S)^2}}\\
SAE_{even} &= \sqrt{S}\\
(\#eq:sae-even)
\end{aligned}
\end{equation}

@@

```{r}
sem_cap <- paste0(
  "Structural Equation Model linking initial model parameters to resulting
  temporal stability of biomass. (a) Arrows display positive (continuous) and
  negative (dashed) standardised effects. Initial parameters were initial
  species richness, initial connectance, Prey-Predator Mass Ratio (PPMR), hill
  exponent, variance of stochastic mortality rates (Env. stochasticity), and
  species response diversity. Only the standardised coefficients whose absolute
  values are superior to 0.1 are displayed. A more complete SEM containing
  further partition of Statistical Averaging Effect and Compensatory effects is
  displayed in Fig. SXX. (b) Total effect of food-web structure on community stability,
  asynchrony, and population stability. The total effects were derived from the
  structural equation model. (c) Partial relationships derived from the SEM, points
  displaying the partial standardized residuals while lines displays the linear
  relationship derived from the SEM.")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here("report", "figures", "p_sem_fig.pdf"))
```

@@

We found that food-web structure had significant effect on temporal population
stability such as average trophic level and type III functional response (i.e.
hill exponent = 3) increase population stability (resp. $r_\delta$= `r rpt` and
`r rph`, $r_\delta$ being the standardized coefficient, Fig. \@ref(fig:sem)a).
In turn, the variance of environmental stochasticity had a strong negative effect
on population stability ($r_\delta$= `r rpe`).

We further found that food-web structure had a strong impact on statistical
averaging effects, which raises from independent species fluctuations weighted
the evenness of their variance. As expected the mathematical relationship (eq.
\@ref(eq:sae-even)), species richness was a main positive driver of statistical
averaging effect ($r_\delta$= `r rsr`, Fig. \@ref(fig:sem)a), thereby increasing
asynchrony. Average trophic level and average interaction strength had strong
negative effects on statistical averaging effects (resp. $r_\delta$= `r rst` and
`r rsi`), thereby downplaying the portfolio effect raised from species richness.
In turn, we found that average interaction strength had a small positive effect
on compensatory dynamics ($r_\delta$= `r rci`, Fig. \@ref(fig:sem)). Species
response diversity had a strong positive effect on compensatory dynamics
($r_\delta$= `r rcd`).

@@

Population stability had a positive effect twice higher than asynchrony on the
temporal stability of community (resp. $r_\delta$= `r rscp` and `r rsca`). In
turn, asynchrony was equally driven by statistical averaging effects and
compensatory dynamics (resp. $r_\delta$= `r rac` and `r ras`, Fig.
\@ref(fig:sem)a).

```{r, eval = FALSE}
semtot_cap <- "Total effect of food-web structure on community stability,
asynchrony, and population stability. The total effects were derived from the
structural equation model."
```

```{r semtot, eval = FALSE, fig.cap = semtot_cap, fig.id = "semtot"}
knitr::include_graphics(here::here("report", "figures", "p_sem_tot_stab.png"))
```

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and response diversity among
species had by far the most important effect on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsdiv`, Fig. \@ref(fig:sem)b). Species richness
had a low overall positive total effect on community stability (`r rtsr`)
because its strong total positive effect on asynchrony (`r rtar`) was downplayed
by its total negative effect on population stability (`r rtpr`) and that the
weight of population stability on community stability is twice higher than the
one of asynchrony. Average trophic level, type III functional response and
predator prey bodysize ratio had positive effects on community stability in
total (resp. `r rtst`, `r rtsh`, `r rtsz`, Fig. \@ref(fig:sem)b), driven by
their positive effects on population stability that were weakly counterbalanced
by their negative effects on asynchrony. In turn, connectance and average
interaction strength had low total negative effects on community stability (resp.
`r rtsc` and `r rtsi`, Fig. \@ref(fig:sem)b). While connectance had negative
effects both on population stability and asynchrony, average interaction
strength had a low total positive effect on population stability counterbalanced
by a stronger negative effect on asynchrony.

@@

)
pred_cap <- paste0("Relationship between community stability and species
  richness according to food-web structure and response diversity. a,
  Predictions of the statistical model linking the temporal stability of
  community biomass and species richness. The values chosen for high/low
  response diversity and food-web structure are indicated by the black point in
  (b). Lines display the mean predictions of the model, ribbons display the
  confidence intervals at 95\\% (to recheck). b, Diagrams displaying the slope
  coefficients of the effect of species richness on community stability. Default
  parameter values for the predictions: ", pv_cap, ".") %>%
  str_replace_all("\\s+", " ")
```

@@

Using a more complex linear model (i.e. including interactions), We found strong
interactions between food-web structure and response diversity on the
relationship between community stability and species richness. We found that
high response diversity almost always leads to positive relationships between
community stability and species richness, no matter the values of average
interaction strength, average trophic level or connectance (Fig.
\@ref(fig:pred)). Furthermore, we found that higher response diversity increased
community stability in average (Fig. \@ref(fig:pred)a). Surprisingly, despite
the relatively small effects of food-web structure on overall community
stability (Fig. \@ref(fig:sem)c \@ref(fig:pred)a), we found that response
diversity dramatically change the shape of the community stability-species
richness relationship from positive to negative when response diversity is low
(Fig. \@ref(fig:pred)b).

---
#We found that
#environmental stochasticity decrease overall community stability but has only
#little effect on the shape of the community stability-species richness
#relationship (Fig. \@ref(fig:pred)a)
---

@@

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, confirming previous results in
empirical fish communities [@danet_species_2021], which contrasts with empirical
data in birds and bats assemblages that reported a equivalent influence of
population stability and asynchrony on community stability
[@olivier_urbanization_2020]. Although the number of empirical and theoretical
studies is still limited to compare the respective contribution of asynchrony and
population stability to community stability in competitive assemblages and
food-webs, we suggest two non-exclusive and complementary explanations that
might drive a lower contribution of asynchrony in food-webs compared to
competitive assemblages. First, the range of bodymasses is much higher in
food-webs, which typically include species that differ in bodymass by several order of
magnitude [@brose_predator_2019], which in turn creates highly uneven biomass
distribution, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits the raise of high portfolio effects and
compensatory dynamics in species rich communities [@zhao_biodiversity_2022].
Second, the trophic interactions characterising food-webs create strong
interdependence among species, which in turn limits the raise of independent
fluctuations (i.e. portfolio effects, but rise compensatory dynamics). The question of
which mechanisms and community features drive differences in the
contribution of asynchrony and population stability in food-webs and competitive
assemblages is still a open question that might be crucial to accurately
address diagnostics and conservation issues. On top of food-web structure,
we found that environmental stochasticity and environmental correlation among
species were the main drivers of community stability, suggesting that stressors
are the main drivers of temporal stability. This result is in accordance with previous
studies showing strong effects on environmental stochasticity
[@vasseur_environmental_2007] and the loss of response diversity decrease
community stability in tri-trophic food-webs [@ives_stability_2000;
@thebault_trophic_2005].

We further found that average trophic level increased average population
stability, confirming previous theoreical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_2023] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators are themselves more stable due
to lower mortality rates with in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
preys [@shanafelt_stability_2018, we can test that?]. In turn, we found that
average trophic level decreases strongly asynchrony through a negative effect on
statistical averaging effect. Further analysis showed that average trophic level
limited statistical averaging effect by decreasing the evenness of biomass
variability. It is linked to the model assumption that species of higher trophic
level have a lower overall mortality rates, similarly to previous studies
[@vasseur_environmental_2007], thereby increasing average population stability and simultaneously
synchronising populations. The fact that Predator Size Mass Ratio displays the same
effect, thereby building food-webs with bigger species having lower mortality
rates, further lends toward this hypothesis. Furthermore, average omnivory was
strongly correlated with average trophic level, suggesting that average omnivory
also stabilise population stability, in line with previous theoretical results
in small modules [@mccann_reevaluating_1997]. Our results further suggest that
higher average trophic level results in lower compensatory dynamics raising from
species interactions. It might be explained by the fact that average trophic
level is strongly associated with average omnivory, which might tamper
synchronized predator-prey oscillations that can raise compensatory dynamics
[@mccann_reevaluating_1997].

We however found no big effect of species richness on population stability,
which has been reported to be negative, positive or neutral in previous studies
[@ruijven_contrasting_2007; @houlahan_negative_2018]. [interpretation?].
However, species richness had a strong positive on asynchrony, by raising
statistical averaging effects. This result is common in empirical systems
[@tilman_biodiversity_2006; @olivier_urbanization_2020], but strongly depends on
the presence of response diversity [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022]. Connectance, quantifying
the generalism in food-webs, had negative effects on both asynchrony and population
stability. We further found that higher connectance leaded to less even biomass
variability, thereby limiting the raise of the portfolio effects, and leaded to
lower compensatory dynamics due to species interactions. This result is in
accordance with empirical findings that a generalist predator can indeed
synchronises the dynamics of a large number of preys
[@raimondo_interspecific_2004]. [No idea about connectance decreasing population
stability]. However, the overall effect of connectance on overall stability was
quite low, similarly to previous empirical findings [@danet_species_2021], but
in apparent disagreement with the predominant role of connectance on stability
in theoretical studies [complexity-stability paradox, @may_will_1972].

---
#In turn,
#we found that species richness magnified the effect of the loss of response
#diversity on the compensatory dynamics due to environment. Our interpretation is
#that higher species richness synchronizes more species fluctuations due to
#environmental mortality when species are not independent because each added
#species adds (S - 1, S being richness) positive covariance terms in the
#stochastic mortality rates, thereby increasing the overall synchrony in
#mortality rates when cross correlation are not null. This assumption that each
#species are equally correlated to each other seems unlikely to be true but as
#far we know, there is compelling evidence yet for alternative correlation
#structures.
---

Food-web dynamical features, the type of functional response (from II to III) and average
interaction strength, also had an effect on community stability. Our results
confirm the stabilising effect of the functional response of type III on
populations [@uszko_effects_2017; @brose_allometric_2006]. Compared to type II,
functional response of type III models a decrease of the feeding rates of
consumers when their preys have a low biomass which leads to a less dramatic
fluctuations of prey biomass. Furthermore, we found that functional response of
type III decreased asynchrony by decreasing compensatory dynamics due to species
interactions. This effect might be due to prey switching of consumers when one of
them reach low abundance [@mccann_diversitystability_2000]. Average interaction
strength, as connectance, had overall little effect on community stability.
Average interaction strength was found to decrease asynchrony by decreasing the
evenness of biomass variability, thereby decreasing portfolio effects,
despite having a small positive effect on compensatory dynamics. In turn, we
surprisingly found that average interaction strength had little effects on
population stability, which contrasts with the importance of this parameter in
complexity-stability relationship [@may_will_1972].

@@

We found that environmental stochasticity does not affect the shape of the
community stability - richness but rather control the overall community
stability (i.e. the intercept of the relation). At the opposite, variation in
response diversity had a strong impact on the relationship between stability and
richness. The reason is that the strength of environmental stochasticity mainly
affect population stability and less asynchrony, the latter being the driver of
species-richness relationships [@ives_stability_2000; @thebault_trophic_2005;
@thibaut_understanding_2013]. Average interaction strength and connectance that
were found to have little effects on overall stability, were found to have a
strong effect on stability-richness relationships. This result is in accordance
with the complexity-stability theory that predicts that community stability
(i.e. measured as the dominant eigen value of the jacobian matrix) should
decrease as the community is species diverse, has higher connectance, and
higher interaction strength [@may_will_1972]. However, we found that the
negative effects of complexity on stability-richness relationship are cancelled
when response diversity is high. A interpretation is that environmental
stochasticity itself does not override the negative complexity-stability
relationships, but it the presence of response diversity and resulting
asynchrony in species fluctuations which overrides the negative effects of
complexity on stability-richness relationships. Response diversity might then be one
mechanism that explains why so much positive relationship between stability and
species richness are found in empirical settings while food-web theoretical
model that do not often include environmental stochasticity and response
diversity finds negative complexity stability relationships
[@ives_stability_2007].
